Read quality check

### 2019-02-21 writing/qual-check.Rmd
Set up for clustermq, require patched version of dada2 which allows to specify fastq quality format, add comments, include full ITS in region extraction

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
datasets <- readd(datasets)
# xdata <- readd(itsxtrim_.its1.lr5.)
# pos <- readd(positions_pb500002A1f)
reads <- readd(qstats_join)
```

@@

## Minimum quality score
```{r}
ggplot(reads, aes(x = minq,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Minimum quality score",
                     # trans = reverselog_trans(10),
                     # limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

@@

ggplot(reads, aes(x = erate,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

## Expected number of errors
```{r}
ggplot(reads, aes(x = eexp,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

@@

```{r}
ggplot(reads, aes(x = p.noerr,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expectation of being error-free",
                     # trans = log_trans(10),
                     limits = c(0, 1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

@@

```{r}
ggplot(reads, aes(x = length,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_log10(name = "Length (bp)") +
  scale_y_continuous(name = "Fraction Greater than length")

```

@@

dada.counts
```
```{r eval = FALSE}
s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         pb.reads = .)
s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         ion.reads = .)

@@

IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`

>>>>

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
```

```{r include=FALSE}
datasets <- readd(datasets)
# xdata <- readd(itsxtrim_.its1.lr5.)
# pos <- readd(positions_pb500002A1f)
reads <- readd(qstats_join) %>%
  mutate_at("Region", factor,
            levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
  mutate_at("Dataset", factor,
            levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
  group_by(Dataset, Plate, Region)
reads_subset <- reads %>%
  do(if (nrow(.) > 10000) sample_n(., 10000) else .)
summarize(reads, n = n())
```

@@

## Minimum quality score
```{r}
ggplot(reads_subset, aes(x = minq,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Minimum quality score",
                     # trans = reverselog_trans(10),
                     # limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

@@

ggplot(reads_subset, aes(x = erate,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

## Expected number of errors
```{r}
ggplot(reads_subset, aes(x = eexp,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

@@

```{r}
ggplot(reads_subset, aes(x = p.noerr,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expectation of being error-free",
                     # trans = log_trans(10),
                     limits = c(0, 1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

@@

```{r}
ggplot(reads_subset, aes(x = length,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_log10(name = "Length (bp)",
                limits = c(50, NA),
                oob = squish) +
  scale_y_continuous(name = "Fraction Greater than length") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

## Read length density
```{r}
ggplot(reads_subset, aes(x = length,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(stat = "density") +
  scale_x_log10(name = "Length (bp)",
                limits = c(50, NA),
                oob = squish) +
  scale_y_continuous(name = "count") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete() +
  theme_bw()
```

@@

dada.counts
```

## ASV correspondences

```{r eval = FALSE}
s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         pb.reads = .)
s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         ion.reads = .)

@@

- IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
- IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
- PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
- PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`
